---
title: "Analysis of Immigration in European Union"
author: "Samuel Mokran"
date: '2024'
output: html_document
runtime: shiny
#file: DAR2_projekt_v4_RMarkdownShiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This R Markdown document is made interactive using Shiny. This project focuses on the migration data from Eurostat. The data is loaded and cleaned, then a summary is provided. 

```{r, echo = F, results = 'hide', message=FALSE}
# load libraries
library(readr) # for loading data
library(tidyverse) # data manipulation
# library(lubridate) # used but without loading, with "::"
```

```{r, echo = F, results = 'hide', message=FALSE}
# load data
data_v6_2 <- read_csv("data/migr_imm1ctz__custom_11840000_linear.csv")
```

### Summary of data
```{r, echo = F}
data_v6_2_clean <- data_v6_2 %>% 
  #select(-c(1, 2, 3, 5, 7)) %>% #replaced
  select(-c("DATAFLOW", "LAST UPDATE", "freq", "agedef", "unit")) %>% 
  #mutate(TIME_PERIOD = as.Date(TIME_PERIOD, format = "%Y")) #replaced
  mutate(TIME_PERIOD = lubridate::ymd(TIME_PERIOD, truncated = 2L)) # instead of as.Date which doesnt work because years are in format 20xx

data_v6_2_clean %>% 
  summary()
```

### Data Insight
average (per year) and total number of immigrants in countries throughout the years
```{r echo=FALSE}
inputPanel(
  textInput("citizenship_1", label = "citizenship:", value = "SK")
)
renderTable({
  req(input$citizenship_1)
  
  data_v6_2_clean %>%
    filter(age == "TOTAL", sex == "T") %>% # duplicate observations
    group_by(geo) %>% # (., TIME_PERIOD)
    summarise(mean_immigrants = format(mean(OBS_VALUE), scientific = FALSE), total_immigrants = sum(OBS_VALUE)) %>%  # scientific format off 
    filter(geo == input$citizenship_1)
})
```

# Data Visualization

choose citizenship and find top 5 destination countries where citizens migrate to
```{r, echo=FALSE}
inputPanel(
  textInput("citizenship", label = "citizenship:", value = "SK"),
  
  sliderInput("citizen_to_top_N", label = "number of top destinations:",
              min = 1, max = 10, value = 5, step = 1)
)
renderPlot({
  req(input$citizenship, input$citizen_to_top_N)
  
  citizen_to_top <- data_v6_2_clean %>%
  filter(citizen == input$citizenship, sex == "T", age == "TOTAL") %>% 
  group_by(geo) %>% 
  summarise(citizen_top_mean = mean(OBS_VALUE)) %>% # sum(OBS_VALUE) vs mean()
  arrange(desc(citizen_top_mean)) #%>% 
  
  top_geos <- citizen_to_top$geo[1:input$citizen_to_top_N]
    
data_v6_2_clean %>% 
  filter(citizen == input$citizenship, geo %in% top_geos, age == "TOTAL", sex == "T") %>%
  group_by(TIME_PERIOD, geo) %>% 
  summarise(mean_immigrants = format(mean(OBS_VALUE), scientific = FALSE), total_immigrants = sum(OBS_VALUE)) %>% 
  mutate(geo = factor(geo, levels = top_geos)) %>% # for order in ggplot
ggplot(aes(x = TIME_PERIOD, y = total_immigrants, color = geo)) + # color = reorder(geo, -total_immigrants)
  geom_line(size = 2) +
  labs(title = "Immigration Trends of Selected EU Citizens" ,
       x = "Year",
       y = "Number of Immigrants",
       color = "Destination Country") + 
  scale_y_continuous(labels = scales::label_number(suffix = " K", scale = 1e-3)) #+ # format y axis
  #scale_color_discrete(name = "Destination Country")
})
```

### Gender Distribution
```{r, echo=FALSE}
renderPlot({
  req(input$citizenship, input$citizen_to_top_N)
  
  citizen_to_top <- data_v6_2_clean %>%
  filter(citizen == input$citizenship, sex == "T", age == "TOTAL") %>% 
  group_by(geo) %>% 
  summarise(citizen_top_mean = mean(OBS_VALUE)) %>% # sum(OBS_VALUE) vs mean()
  arrange(desc(citizen_top_mean)) #%>% 
  
  top_geos <- citizen_to_top$geo[1:input$citizen_to_top_N]
  
  data_v6_2_clean %>% 
    filter(citizen == input$citizenship, geo %in% top_geos, age == "TOTAL", sex != "T") %>% 
    group_by(geo, sex) %>% 
    summarise(mean_immigrants = format(mean(OBS_VALUE), scientific = FALSE), total_immigrants = sum(OBS_VALUE)) %>% 
    mutate(geo = factor(geo, levels = top_geos)) %>% # for order in ggplot
    ggplot(aes(x = factor(sex), y = total_immigrants, fill = geo)) +
    geom_bar(stat = "identity") + coord_flip() +
    labs(title = "Immigration Trends of Selected EU Citizens",
         x = "Sex",
         y = "Number of Immigrants",
         fill = "Destination Country") + # legend for geo v2
    scale_y_continuous(labels = scales::label_number(suffix = " K", scale = 1e-3)) #+ # format thousands
    #scale_fill_discrete(name = "Destination Country") # legend for geo v1
})
```

### Age Distribution
```{r, echo=FALSE}
renderPlot({
  req(input$citizenship, input$citizen_to_top_N)
  
  citizen_to_top <- data_v6_2_clean %>%
  filter(citizen == input$citizenship, sex == "T", age == "TOTAL") %>% 
  group_by(geo) %>% 
  summarise(citizen_top_mean = mean(OBS_VALUE)) %>% # sum(OBS_VALUE) vs mean()
  arrange(desc(citizen_top_mean)) #%>% 
  
  top_geos <- citizen_to_top$geo[1:input$citizen_to_top_N]

  # summarize the data by age group and country
  summary_data_age <- data_v6_2_clean %>% 
    filter(citizen == input$citizenship, geo %in% top_geos, sex != "T") %>% 
    filter(!age %in% c("TOTAL", "UNK", "Y_GE100", "Y_GE65", "Y_GE85", "Y_LT15", "Y_LT5", "Y15-64")) %>% 
    group_by(geo, age) %>% 
    summarise(total_immigrants = sum(OBS_VALUE)) %>% 
    mutate(geo = factor(geo, levels = top_geos)) # for order in ggplot
  
  # bar plot
  ggplot(summary_data_age, aes(x = age, y = total_immigrants, fill = geo)) +
    geom_bar(stat = "identity", position = "dodge") + 
    labs(title = "Total Number of Immigrants by Age Group and Country",
         x = "Age Group",
         y = "Number of Immigrants",
         fill = "Destination Country") + # legend for geo v2 
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    scale_y_continuous(labels = scales::label_number(suffix = " K", scale = 1e-3)) #+ # format thousands
    #scale_fill_discrete(name = "Destination Country") # legend for geo v1
})
```

## END




